---
title: "comparative structure analysis"
format: pdf
editor: visual
---

## Comparative structure analysis

here we run through a complete "pipline" of structure analysis that begins with a single sequence identifier and ends in a PCA analysis.

```{r}
library(bio3d)
```

## step 1. retrive a sequence for the protein we are interested in . we will take ADK "1ake_A"

```{r}
id <- "1ake_A"
aa <- get.seq(id)
aa
```

## step 2.

run a BLAST search of the pdb for all the related sequences to our input `aa`

```{r}
blast <- blast.pdb(aa)
```

```{r}
hits <- plot(blast)
```

## step3. download structure

these are top hits ie all structures in the pdb data related to our input sequence

```{r}
hits$pdb.id
```

```{r}
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

## setp 4. allign the superpose

```{r}
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")

```

```{r}
pdbs
```

## step 5. PCA

lets use our old friend to make sense of

```{r}
pc <- pca(pdbs)
plot(pc)
```
