---
title: "lab8.0"
format: pdf
editor: visual
---

## Quarto

```{r}
head(mtcars)
```

```{r}
apply(mtcars, 2, sd)
```

it is clear "disp" and "hp" have the highest mean values and the highest standard deviation here. they will likely dominate any analysis i do on this dataset. lets see

```{r}
pc.noscale <- prcomp(mtcars, scale=FALSE)
pc.scale <- prcomp(mtcars, scale=TRUE)
```

```{r}
biplot(pc.noscale)
```

```{r}
pc.noscale$rotation[,1]
```

plot the loadings

```{r}
library(ggplot2)
r1 <- as.data.frame(pc.noscale$rotation)
r1$names <- rownames(pc.noscale$rotation)


ggplot(r1) + 
  aes(PC1, names) + 
  geom_col()
```

```{r}
library(ggplot2)
r1 <- as.data.frame(pc.scale$rotation)
r1$names <- rownames(pc.scale$rotation)


ggplot(r1) + 
  aes(PC1, names) + 
  geom_col()
```

> **Take Home**: generally we always want to set 'scale=TRUE' when we do this type of analysis to avoid our analysis being dominated by individual variables with the largest variance just due to their unit of measurement

# FNA breast cancer data

load the data into R.
